### 2020-05-31 report/sections/evalconcl/eval/benchmarks.tex
Report: benchmarks

\item The entry point to the \trole{Svr} logic is parameterised by the
number of messages, $n$, configured through an environment variable
passed through the command line.

\item We use the transpiled JavaScript versions of all \trole{Svr}
for the experiments.

@@

We run our experiments using three implementations of the Ping Pong protocol:

@@

\subparagraph{mpst:}
The \texttt{mpst} implementation follows the MPST framework and
uses the APIs generated from \fancyname{SessionTS}, so it enjoys the
communication safety guarantees from our methodology.

\subsection{Execution Pattern}
We compare the execution patterns of exchanging 10,000 Ping-Pongs
throughout 20 repeated experiments across the three implementations.
We visualise the elapsed time with respect to the number of \tmsg{PING}s
received in \cref{fig:execution}.

\begin{figure}[!h]
\centering
\begin{subfigure}[b]{0.3\textwidth}
\centering
\includegraphics[width=\textwidth]{execbare10000}
\caption{\texttt{bare}}
\label{fig:executionbare}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.3\textwidth}
\centering
\includegraphics[width=\textwidth]{execbaresafe10000}
\caption{\texttt{bare_safe}}
\label{fig:executionbaresafe}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.3\textwidth}
\centering
\includegraphics[width=\textwidth]{execmpst10000}
\caption{\texttt{mpst}}
\label{fig:executionmpst}
\end{subfigure}
\captionof{figure}{Comparison of Execution Pattern for 10,000 Ping-Pongs}
\label{fig:execution}
\end{figure}

\subsection{Overhead}
We compare the total execution time (\textit{Exe. Time}) 
and execution time per round trip (\textit{Exe. Time / Ping-Pong}) -- 
averaged over 20 repeated experiments -- across the three implementations,
for $n \in \{10^2, 10^3, 10^4\}$.
We summarise the results in \cref{table:overhead}.

\renewcommand{\arraystretch}{1.5}
\begin{table}[!h]
\centering
\begin{tabular}{||c||c|c|c||c|c|c||}
\hline
\multirow{2}{*}{$n$} & 
\multicolumn{3}{c||}{Exe. Time} & 
\multicolumn{3}{c||}{Exe. Time / Ping-Pong} \\
\cline{2-7}
 & \texttt{bare} & \texttt{bare_safe} & \texttt{mpst} 
 & \texttt{bare} & \texttt{bare_safe} & \texttt{mpst} \\
\hline\hline
$10^2$ & 89.64ms & 107.09ms & 186.23ms & 0.90ms & 1.07ms & 1.86ms \\
$10^3$ & 642.92ms & 663.91ms & 1155.48ms & 0.64ms & 0.66ms & 1.16ms \\
$10^4$ & 3542.16ms & 3837.97ms & 7015.25ms & 0.35ms & 0.38ms & 0.70ms \\
\hline
\end{tabular}
\captionof{table}{Comparison of Execution Time for 
100, 1,000 and 10,000 Ping-Pongs}
\label{table:overhead}
\end{table}
\renewcommand{\arraystretch}{1}

@@

\begin{figure}[!h]
\centering
\includegraphics[width=\textwidth]{timepermsg}
\captionof{figure}{Comparing Average Time per Ping-Pong 
Across Implementations}
\label{fig:timepermsg}
\end{figure}

The \texttt{mpst} implementation records greater round trip times compared
to both \texttt{bare} and \texttt{bare_safe} variants. This is expected,
as \dots

>>>>

\item \trole{Svr} logic is parameterised by the number of round trips,
$n$, configured through an environment variable passed through 
the command line.

\item We use the transpiled JavaScript versions of all \trole{Svr}s
for the experiments.

@@

The benchmark compares three implementations of the Ping Pong protocol:

@@

\subparagraph{mpst:}
The \texttt{mpst} implementation uses the APIs generated from 
\fancyname{SessionTS}, so it enjoys the
communication safety guarantees from our methodology.

\subsection{Execution Pattern}
\label{section:execpattern}

We compare the execution patterns of exchanging 10,000 Ping-Pongs
throughout 20 repeated experiments across the three implementations.
We visualise the elapsed time with respect to the number of \tmsg{PING}s
received in \cref{fig:execution}.

\begin{figure}[!ht]
\centering
\begin{subfigure}[b]{0.3\textwidth}
\centering
\includegraphics[width=\textwidth]{execbare10000}
\caption{\texttt{bare}}
\label{fig:executionbare}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.3\textwidth}
\centering
\includegraphics[width=\textwidth]{execbaresafe10000}
\caption{\texttt{bare_safe}}
\label{fig:executionbaresafe}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.3\textwidth}
\centering
\includegraphics[width=\textwidth]{execmpst10000}
\caption{\texttt{mpst}}
\label{fig:executionmpst}
\end{subfigure}
\captionof{figure}{Comparison of Execution Pattern for 10,000 Ping-Pongs}
\label{fig:execution}
\end{figure}

Relative to the two \texttt{bare} implementations, the
\texttt{mpst} implementation shows more consistency.

\subsection{Overhead}
We compare the total execution time (\textit{Exec. Time}) 
and execution time per round trip (\textit{Exec. Time / Ping-Pong}) -- 
averaged over 20 repeated experiments -- across the three implementations,
for $n \in \{10^2, 10^3, 10^4\}$.
We summarise the results in \cref{table:overhead}.

\renewcommand{\arraystretch}{1.5}
\begin{table}[!ht]
\centering
\begin{tabular}{||c||c|c|c||c|c|c||}
\hline
\multirow{2}{*}{$n$} & 
\multicolumn{3}{c||}{Exec. Time} & 
\multicolumn{3}{c||}{Exec. Time / Ping-Pong} \\
\cline{2-7}
 & \texttt{bare} & \texttt{bare_safe} & \texttt{mpst} 
 & \texttt{bare} & \texttt{bare_safe} & \texttt{mpst} \\
\hline\hline
$10^2$ & 89.64ms & 107.09ms & 186.23ms & 0.90ms & 1.07ms & 1.86ms \\
$10^3$ & 642.92ms & 663.91ms & 1155.48ms & 0.64ms & 0.66ms & 1.16ms \\
$10^4$ & 3542.16ms & 3837.97ms & 7015.25ms & 0.35ms & 0.38ms & 0.70ms \\
\hline
\end{tabular}
\captionof{table}{Comparison of Execution Time for 
100, 1,000 and 10,000 Ping-Pongs}
\label{table:overhead}
\end{table}
\renewcommand{\arraystretch}{1}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\textwidth]{timepermsg}
\captionof{figure}{Comparing Average Time per Ping-Pong 
Across Implementations}
\label{fig:timepermsg}
\end{figure}

The \texttt{mpst} implementation records greater round trip times compared
to both \texttt{bare} and \texttt{bare_safe} variants. This is expected,
as the runtime for \textit{each} role performs additional logic during
\textit{both} the sending and receiving of messages. 
For example, receiving a message involves performing
(albeit $O(1)$ time complexity) operations on the message queue and
handler queue. As for browser-side implementations, every EFSM transition
invokes an updated \texttt{render()} on the VDOM, which requires
reconciliation internally by React to update the browser DOM accordingly.

We also observe that the (round trip time) performance gap between 
the \texttt{mpst} implementation and the two 
\texttt{bare} implementations narrows as the number of round trips 
increase. This suggests that the session joining logic in our
implementation yields greater overhead than the additional processing
logic injected during communication, which is consistent with 
our findings in \cref{section:execpattern}.

We interpret the overhead as a trade-off between maximising performance and
maximising the \textbf{static} communication safety guarantees for
web applications. In particular, the event-driven nature of 
browser-side logic makes it highly challenging to guarantee 
communication safety, as having ``active'' channel actions on the DOM
allows linearity to be violated by the user, even if one verifies
linear channel usage in the source code. The \texttt{mpst} implementation
properly addresses the problem by statically guaranteeing,
by construction, that the UI component rendered on the DOM will only
contain channel actions permitted at that state in the EFSM execution.

One may argue that the \texttt{bare_safe} implementation also
provides said guarantee but adds negligible overhead. 
By inspecting the workaround in the \texttt{bare_safe} implementation
(\cref{lst:workaround}),
it is clear that this does not scale for more complicated protocols with
larger EFSMs.

\begin{figure}[!ht]
\begin{lstlisting}[language=javascript, tabsize=2]
click() {
	this.state.ws?.send(JSON.stringify({
		label: 'PING',
		payload: [this.context.count],
	}));
	this.setState({ visible: false });
}

render() {
	return this.state.visible 
		? <button
		   ref={this.state.button}
		   onClick={this.click.bind(this)}
		   >Ping</button>
		: <button>Ping</button>;
}
\end{lstlisting}
\captionof{lstlisting}{Preventing user-triggered linearity violation in
\texttt{bare_safe} Ping Pong}
\label{lst:workaround}
\end{figure}

The workaround is tailored to the Ping Pong protocol
and toggles the \texttt{visible} flag to hide the button that triggers
a send action when \trole{Client} transitions to a receive state.
In fact, this ``workaround'' precisely generalises to our 
\texttt{mpst} implementation of having some form of wrapper
component that renders the UI component corresponding to the
current EFSM state. Our API generation strategy formally implements
the EFSM for browser-side logic.

We also compare the server-side logic between the \texttt{base} and
\texttt{mpst} implementations in \cref{fig:nodepingpong}, and observe
that our generated APIs (\cref{lst:nodepingpongmpst}) 
allow the developer to focus on the
implementation detail, whilst the naive bare implementation 
(\cref{lst:nodepingpongbare}) chooses to interleave communication logic
with application logic, which arguably contributes towards a hidden
source of bugs found when implementing more complex protocols.

\begin{figure}[!ht]
\centering
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
socket.onmessage = ({ data }) => {
	const { label, payload } = JSON.parse(data.toString());
	if (label === 'PING') {
		let count: number = payload[0];
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			socket.send(JSON.stringify({
				label: 'BYE',
				payload: [count],
			}));
		} else {
			socket.send(JSON.stringify({
				label: 'PONG',
				payload: [count],
			}));
		}
	} else {
		throw new Error(`Unrecognised label: ${label}`)
	}
}
\end{lstlisting}
\caption{\texttt{bare}}
\label{lst:nodepingpongbare}
\end{subfigure}
\hfill
\begin{subfigure}[b]{\textwidth}
\centering
\begin{lstlisting}[language=javascript,tabsize=2]
// NOTE: `Implementation' namespace omitted for brevity.

const logic = new Initial({
	PING: (count) => {
		console.timeLog(LABEL, ++count);
		if (count === MSGS) {
			return new S16([Labels.S16.BYE, [count], new Terminal()]);
		} else {
			return new S16([Labels.S16.PONG, [count], logic]);
		}
 	}
});
\end{lstlisting}
\caption{\texttt{mpst}}
\label{lst:nodepingpongmpst}
\end{subfigure}
\captionof{figure}{Comparison of \trole{Svr} implementations in benchmark}
\label{fig:nodepingpong}
\end{figure}
